### 2023-08-31 sage/template.qmd

---
title: Sage Template
format:
  sage-pdf:
    keep-tex: true  
  sage-html: default
author:
  - name: Sarah Malloc
    affiliations:
      - name: An Organization
        department: The Department
        address: 1 Main St
        city: Boston
        country: USA
        postal-code: 02210-1022
      - A second affilication
    orcid: 0000-0000-0000-0000
    email: sm@example.org
    url: https://example.org/
  - name: Eliza Dealloc
    affiliations:
      - Another Affiliation
abstract: |
  This document is a template demonstrating the Sage format.
keywords: [template, demo]
bibliography: bibliography.bib  
---

## Introduction {#sec-intro}

*TODO* Create a template that demonstrates the appearance, formatting, layout, and functionality of your format. Learn more about journal formats at <https://quarto.org/docs/journals/>.