### 2024-06-10 tandf/Response to reviewers jg revised.qmd
updated files from jg

---
title: "Response to reviewers"
format: pdf
---

## Reviewer 1

*However, the introduction of the current manuscript should be rewritten. The authors used lots of sentences to describe the background of social media use in China to introduce the research questions. This is unhelpful. It is suggested to examine the related theories to answer, why the phenomenon of social media use in Mainland China may be different from others.*

We introduce the context of China primarily in the Literature Review and Theory section, so we are assuming this is the section to which the reviewer is referring. We agree that the China section is perhaps too long. We have trimmed it and added linkages to relevant theories.

## Reviewer 2

*1.	In the abstract, there is more information that needs to be added: a clear aim of this research and a few words about the sample size.*

Abstract rewritten and sample size added.

*2.	What is FOMO in abstract? Needs to be clarified here to be understood by the readers.*

Definition added.

*3.	References need to be added in the Introduction section. Also, I suggest merging the introduction section with the literature review and theory section.*

References added. We welcome additional feedback on how to merge the introduction and literature review sections but we believe after deleting the first paragraph of the lit review section, there is a clearer delineation between the two sections.

*4.	In Literature Review and Theory subtitle, the first sentence needs a reference.*

The sentence is deleted as per the comment to point 5 below.

*5.	The second and third sentence in the Literature Review and Theory section (first paragraph) can be moved in the end of the introduction section or can be omitted.*

The sentences have been deleted, additionally, the first sentence read awkwardly without the following sentences and did not add significantly to the section, so we also chose to delete that sentence.

*6.	Reference is needed in the first sentence of the second paragraph in the Literature Review and Theory section.*

The references for this claim was located at the end of the paragraph where the arguments of the authors were described in detail. Though, we can see how it would be confusing. We moved the references to the end of the first sentence.

*7.	Nothing in the literature review about social media addiction. I suggest to see: Social isolation, social support and their relationship with smartphone addiction; examples: The relationship between addiction to smartphone usage and depression among adults: A crosssectional study, Smartphone addiction among university students in the light of some variables, and Examining the effects of motives and gender differences on smartphone addiction.*

We agree that smartphone addiction, depression, and feelings of loneliness and a sense of isolation may be related.These works have been added to the literature review.

*8.	I believe that the testable hypotheses are clear but I suggest to be more in a direct way. Why use “weak or possibly” for example? Why “a strongly”? The results will explain these.*

Fair point. Qualifying words have been removed.

*9.	Also, a clear research question need to be stated in the end of the introduction section.*

We re-wrote the second half of the introduction to clarify our research question and contribution.

*10.	How can the researchers explain that the sample represent ALL China? We know the cultural diversity in a “BIG” China.*

We added more details about the sampling strategy in the introduction of the data and measurement section to better explain the strengths and weaknesses of our data when it comes to their external validity.

*11.	Need more explanation about the sample. Researchers stated it is random sample. How? Clarification is needed.*

See point 10.

*12.	I suggest that the method section needs to start with sample.*

The first paragraph of the methods section now addresses the sample in more detail.

*13.	No need for the margin in the end of page 5. This information is vital and essential procedures. Need to be moved in the body of the research.*

We are not sure to exactly what the reviewer is referring here. We think the reference is to the footnote at the end of page 5, and there's a typo in the reviewer's comment ("margin" should be "footnote"). We assumed this was the case. If so, we decided to leave the footnote as is. Because we added more description of the sample in the main body of the text, and the footnote draws attention to the fact that the adjustments we made were not related to the indicators used in the present studies, we thought it would unnecessarily interrupt the flow of the text.

*14.	Validity and reliability are needed for the measurements.*

We included the alpha statistic for the dependent variable; it is high at 0.86. We also added the reliability estimate for social media use. It was not high ($\alpha = 0.31$) but we moved forward with the index because the face validity is pretty clear, those who claim to use social media generally more often, claim to read political news more often, and claim to use it more than they did in the past, are likely to be more frequent social media users than those who do not do these things. Also we added confirmation to the measurement description that the three social media use items were all positively correlated with each other (p < 0.0001 for all pairwise relationships).

*15.	Figure ?? and Table ?? are  noticed. Need to number the figures and tables.*

This may be a font issue, as our version of the PDF does not contain any missing figure or table numbers, nor does the LaTeX file. Happy to help troubleshoot further but everything on our end looks complete.