### 2024-08-27 tandf/Response to reviewers second revision jg edits.qmd
Second response to reviewers

---
title: "Response to reviewers - second revision"
format: pdf
---

## Reviewer 2

*1. Timeliness of Data: The research is based on data collected in 2015, a time when the prevalence of mobile internet and social media platforms might differ significantly from the present landscape. The authors utilized a survey conducted from November 25 to December 2, 2015, as mentioned in the document: "We employed Qualtrics to collect the data...from November 25 to December 2, 2015." Given the rapid evolution of digital communication, the applicability of findings based on a decade-old dataset should be reconsidered.*

We agree that the digital environment changes rapidly, that our data are getting dated, and that we needed to make a case for why we are confident that our theory and results are still relevant today. We make the case that our estimates are actually conservative ones because the new developments in the digital environment should increase the strength of the relationship between social media use and social isolationism. As such, we added to the manuscript a three-pronged defense of the age of the data: 1) Based on our theoretical framework, the development of new social media technologies should only exacerbate the differences between urban and rural variation in the relationship between social media use and feelings of isolation. The new technologies make it even more apparent to rural citizens that they may be missing out - the rise of short video content, improved algorithms that better target users with content and tighten their social networks, better content creation tools, etc. The modern digital environment relative to the one at the time our data were collected should heighten the fear of missing out making our estimates conservative ones. 2) Our questions of digital media consumption are general enough to not be context and time dependent. They are not aimed at specific technologies that are no longer in use. 3) The urbanization of China has increased since the data were collected. This makes citizens living in rural areas increasingly separated from many of their family and friends forced to move to cities for jobs. As such, here too, we expect the divide is only exacerbating the problem. While it is true that more people are moving to cities so they will not have the fear of missing out, those people who are left are less likely to have family and friends nearby.

See footnote 3 for our discussion of this issue.

*2. Mismatch Between Literature Review and Data Collection: The literature review incorporates studies from around 2020, such as the reference to "Valkenburg et al., 2022," yet the data used for analysis is from 2015. This discrepancy may affect the validity of conclusions drawn from contemporary theoretical perspectives when applied to outdated data.*

We are happy to revise our literature review section in ways that the reviewer feels would strengthen the paper but we are a bit unsure of how to proceed given two previous comments by reviewers that we add to the literature review section. While we acknowledge that the previous recommendation and this recommendation are not precisely mutually exclusive, it does suggest that if we remove some references here we may be working against the recommendations of the previous reviewers.

Suggestions as to how to move forward would be welcome here.

> Previous reviewer 1 comment: It is suggested to examine the related theories to answer, why the phenomenon of social media use in Mainland China may be different from others.

> Previous reviewer 2 comment: Nothing in the literature review about social media addiction. I suggest to see: "Social isolation, social support and their relationship with smartphone addiction", "The relationship between addiction to smartphone usage and depression among adults: A crosssectional study", "Smartphone addiction among university students in the light of some variables," and "Examining the effects of motives and gender differences on smartphone addiction."

*3. Rationale for Variable Selection: The inclusion of variables like "Chinese Communist Party membership (CCP)" is noted, as stated: "For those we begin by fitting an additive model of our index of social isolationism as function of each of our primary independent variables...CCP Member." However, the introduction does not discuss the relevance of this variable, leaving unclear its contribution to the study.*

A reasonable point. We have added some discussion of the reason this variable is included in the final paragraph of the Data, Measurement and Description section.

*4. Depth of Discussion: The discussion section lacks depth, particularly in expanding upon the implications of the results. For example, the document mentions: "our results cast some doubt on theories that suggest it is the negative content..." but does not elaborate on how these findings challenge existing theories.*

Fair point. We have substantially increased discussion of the implications and linked it more explicitly back to the literature discussed in the theory section.